Integer Programming Formulations for the Shared Multicast Tree Problem

### 2018-09-23 smt/sec6-exp.tex
Talbe comparing meta algorithm and B&B filled. Added the description of the results.

\textcolor{blue}{
To the best of our knowledge, SMT has not been studied extensively in terms of inexact methods.
However, there is a manifold of a thorough research of the minimum Steiner tree problem, with the possibility of adapting the methods on SMT.
We turn our attention to recently pusblished paper \citet{pajor18} which notains a metaheuristic algorithm for a fast generation of good solution to the Steiner tree problem.
The method is general, and is therefore very suitable for adaptation to our purposes.
The high level description is as follows:}
\begin{algorithm}
\textcolor{blue}{
Create an empty pool $P$ of elite solutions with capacity $\gamma=\left\lceil\sqrt{M/2}\right\rceil$\;
\For {$i\in \{1,\dots,M$\}} {
Generate a new solution $S$ using a constructive algorithm, local search, and randomization\;
\If {$i<\gamma$} {$P\leftarrow P\cup \{S\}$}
\Else {
$\hat{S}\leftarrow\arg\max_{S'\in P} c(S')$\;
\If {$c(\hat{S})> c(S)$}{$P\leftarrow P\cup \{S\}\setminus\{\hat{S}\}$}
$\hat{S}\leftarrow\arg\max_{S'\in P} c(S')$\;
Generate a solution $\tilde{S}$ by combining $S$ with solutions in $P$\;
\If {$c(\hat{S})> c(\tilde{S})$}{$P\leftarrow P\cup \{\tilde{S}\}\setminus\{\hat{S}\}$}
}
}
\Return $\arg\min_{S'\in P} c(S')$\;
 \caption{Outline of the metaheuristic algorithm}
}
\label{alg:meta}
\end{algorithm}
\textcolor{blue}{The construction procedure and local search is the adapted BIP algorithm together with refinements presented in \citet{ivanova16isco}.
The remaining parts of the algorithm are carried out as discussed in \citet{pajor18}.
The randomization is achieved by perturbing the edge costs in the input graph.
The obtained solution $S$ is evaluated with respect to the real (original) edge costs and added to the pool $P$ if the pool capacity is not filled.
If the pool is full and its worst solution $\hat{S}$ is worse than $S$, $\hat{S}$ is replaced by $S$ in $P$.
Then a new solution $\tilde{S}$ is constructed by combining some solution in $P$ with $S$. 
First, we adjust edge costs resulting in a modified input instance, 
and then run the construction and local search yielding $\tilde{S}$.
This new solution is again added to the pool if it is better than some solution already in $P$.
The adjustment of the edges tries to promote edges that appear in both $S$ and a selected solution in $P$. 
In that case, the original cost is preserved.
If an edge is present in only one of the two solutions, its cost is multiplied by a random integer between 100 and 500.
If an edge appears in neither of the solutions, its edge is multiplied by 1000, which makes it very unlikely to appear in the result $\tilde{S}$ of the combination.
}
\begin{table}[]
\centering
\textcolor{blue}{
\begin{tabular}{rrrrrrrrrrr}
\multicolumn{1}{l}{}      & \multicolumn{1}{l}{}      & \multicolumn{3}{c}{5 min}  & \multicolumn{3}{c}{10 min} & \multicolumn{3}{c}{20 min}     \\\hline
\multicolumn{1}{l}{$|V|$} & \multicolumn{1}{l}{$|D|$} & \multicolumn{1}{l}{$\mathcal{X}_1$} & \multicolumn{1}{l}{$\mathcal{F}_1$} & \multicolumn{1}{l}{$\mathcal{M}$} & \multicolumn{1}{l}{$\mathcal{X}_1$} & \multicolumn{1}{l}{$\mathcal{F}_1$} & \multicolumn{1}{l}{$\mathcal{M}$} & \multicolumn{1}{l}{$\mathcal{X}_1$} & \multicolumn{1}{l}{$\mathcal{F}_1$} & \multicolumn{1}{l}{$\mathcal{M}$} \\\hline
20  & 10   &     &   &   &   &    &    &    &    &       \\
22  & 11   &     &   &   &   &    &    &    &    &       \\
24  & 12   &     &   &   &   &    &    &    &    &                                  
\end{tabular}
}
\caption{\textcolor{blue}{Comparison of the metaheuristic algorithm and solving IP models with imposed time-limit.}}
\label{tab:metamod}
\end{table}

\textcolor{blue}{
Tab.\ref{tab:metamod} summarizes the comparison of the metaheuristic algorithm with the ability of IP models to produce feasible solutions when restricted time limit.
Each column represents the average percentage of objective function of the best method in a given instance.
The averages are calculated over all instances of given $V$ and $D$.
}

\textcolor{blue}{
The resulting values reveal that...
}

@@

We compare the progress of the lower bounds obtained by the CG method with the one obtained by B\&B applied to $\mathcal{F}_1$ as a function of running time.
Instances with $|V|=2|D|\in\{24, 26,\ldots, 34\}$ are investigated, with 5 instances for each value of $|V|$.
The results are illustrated in Fig. \ref{fig:cggr}.
Each curve represents the progress of the lower bound within each instance set.
The horizontal axes represent running times ranging from 0 to 20 CPU-minutes.
The vertical axes correspond to the lower bounds as a percentage of $z\left(\text{LP}(\mathcal{F}_1)\right)$, averaged over all instances in the set.

@@

    \begin{subfigure}[b]{0.49\textwidth}
        \includegraphics[width=\textwidth]{lower-bound-32-16}
        \caption{$|V|=32, |D|=16$}
        \label{fig:cggr32-16}
    \end{subfigure}
    \hfill %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
      %(or a blank line to force the subfigure onto a new line)
    \begin{subfigure}[b]{0.49\textwidth}
        \includegraphics[width=\textwidth]{lower-bound-34-17}
        \caption{$|V|=34, |D|=17$}
        \label{fig:cggr34-17}
    \end{subfigure}
  \caption{Comparison of lower bounds obtained by CG and B\&B on model $\mathcal{F}_1$.
	\textcolor{blue}{Each figure depicts a progress of lower bounds for instances with varying number of destination and non-destination nodes.
	The lower bounds are obtained by solving CG scheme applied on model $\mathcal{F}_1$ (black curves), and produced during the course of classic B\&B algorithm (grey curves)
	The tightness of the bounds increases with time, and it can be observed from the graphs that the CG bound becomes tighter at some point.
	We express the quality of lower bounds with respect to the value of simple continuous relaxatin, as indicated by the vertical axes.
	It can also be noted that the lower bound obtained from B\&B grows sharply during the computation of the simplex method in the root of the B\&B tree, 
	which is happening until the curve reaches 100\%.
	After that point the bound grows rather slowly.
	CG on the other hand produces bound that grows faster even after the value of first relaxation in the root is reached.
	}}
  \label{fig:cggr}
\end{figure}

>>>>

\textcolor{blue}{
To the best of our knowledge, SMT has not been studied extensively in terms of inexact methods.
However, there is a manifold of a thorough research of the minimum Steiner tree problem, with the possibility of adapting the methods on SMT.
We turn our attention to recently published paper \citet{pajor18} which contains a metaheuristic algorithm for a fast generation of good solution to the Steiner tree problem.
The method is general, and is therefore very suitable for adaptation to our purposes.  The high level description is as follows:}

\begin{algorithm}
\textcolor{blue}{
Create an empty pool $P$ of elite solutions with capacity $\gamma=\left\lceil\sqrt{M/2}\right\rceil$\;
\For {$i\in \{1,\dots,M$\}} {
Generate a new solution $S$ using a constructive algorithm, local search, and randomization\;
\If {$i<\gamma$} {$P\leftarrow P\cup \{S\}$}
\Else {
$\hat{S}\leftarrow\arg\max_{S'\in P} c(S')$\;
\If {$c(\hat{S})> c(S)$}{$P\leftarrow P\cup \{S\}\setminus\{\hat{S}\}$}
$\hat{S}\leftarrow\arg\max_{S'\in P} c(S')$\;
Generate a solution $\tilde{S}$ by combining $S$ with solutions in $P$\;
\If {$c(\hat{S})> c(\tilde{S})$}{$P\leftarrow P\cup \{\tilde{S}\}\setminus\{\hat{S}\}$}
}
}
\Return $\arg\min_{S'\in P} c(S')$\;
 \caption{Outline of the metaheuristic algorithm}
}
\label{alg:meta}
\end{algorithm}
\textcolor{blue}{The construction procedure and local search is the adapted BIP algorithm together with refinements presented in \citet{ivanova16isco}.
The remaining parts of the algorithm are carried out as discussed in \citet{pajor18}.
The randomization is achieved by perturbing the edge costs in the input graph.
The obtained solution $S$ is evaluated with respect to the real (original) edge costs and added to the pool $P$ if the pool capacity is not filled.
If the pool is full and its worst solution $\hat{S}$ is worse than $S$, $\hat{S}$ is replaced by $S$ in $P$.
Then a new solution $\tilde{S}$ is constructed by combining some solution in $P$ with $S$. 
First, we adjust edge costs resulting in a modified input instance, 
and then run the construction and local search yielding $\tilde{S}$.
This new solution is again added to the pool if it is better than some solution already in $P$.
The adjustment of the edges tries to promote edges that appear in both $S$ and a selected solution in $P$. 
In that case, the original cost is preserved.
If an edge is present in only one of the two solutions, its cost is multiplied by a random integer between 100 and 500.
If an edge appears in neither of the solutions, its edge is multiplied by 1000, which makes it very unlikely to appear in the result $\tilde{S}$ of the combination.
}
\begin{table}[]
\centering
\setlength{\tabcolsep}{6pt} % Default value: 6pt
\renewcommand{\arraystretch}{1.4} % Default value: 1
\textcolor{blue}{
\begin{tabular}{rrrrrrrrrrr}
\multicolumn{1}{l}{}      & \multicolumn{1}{l}{}      & \multicolumn{3}{c}{5 min}  & \multicolumn{3}{c}{10 min} & \multicolumn{3}{c}{20 min}     \\\hline
\multicolumn{1}{l}{$|V|$} & \multicolumn{1}{l}{$|D|$} & \multicolumn{1}{c}{$\mathcal{X}_1$} & \multicolumn{1}{c}{$\mathcal{F}_1$} & \multicolumn{1}{c}{$\mathcal{M}$} & \multicolumn{1}{c}{$\mathcal{X}_1$} & \multicolumn{1}{c}{$\mathcal{F}_1$} & \multicolumn{1}{c}{$\mathcal{M}$} & \multicolumn{1}{c}{$\mathcal{X}_1$} & \multicolumn{1}{c}{$\mathcal{F}_1$} & \multicolumn{1}{c}{$\mathcal{M}$} \\
\hline
18  & 12   &100.0&100.0&101.8&100.0&100.0&100.0&100.0&100.0&100.0\\
21  & 14   &106.5&107.4&101.2&102.7&103.1&100.1&100.9&100.3&100.0\\
24  & 16   &144.2&112.2&101.3&124.0&114.0&100.1&104.3&112.7&100.0\\     
\hline
20  & 10   &100.6&100.0&101.4&100.1&100.0&100.0&100.1&100.0&100.0\\
22  & 11   &102.1&103.2&101.7&101.3&100.7&100.1&100.7&100.2&100.1\\
24  & 12   &106.6&111.7&101.7&105.0&105.5&100.0&103.1&106.2&100.0\\                               
\hline
21  & 7    &100.1&100.0&100.0&100.0&100.0&100.0&100.0&100.0&100.0\\
24  & 8    &102.3&101.4&102.0&102.1&100.7&100.1&101.1&100.1&100.0\\
27  & 9    &127.4&114.0&101.6&111.2&109.0&100.1&108.2&104.1&100.0                                  
\end{tabular}
}
\caption{\textcolor{blue}{Comparison of the metaheuristic algorithm and solving IP models with imposed time-limit.}}
\label{tab:metamod}
\end{table}

\textcolor{blue}{
Tab.\ref{tab:metamod} summarizes the comparison of the metaheuristic algorithm with the ability of IP models to produce feasible solutions when restricted time limit.
Each entry represents the average percentage  of the objective value yielded by the most successful method in a given instance.
For example, if all methods find the optimum in every instance, the three corresponding values are all 100.0.
Thus, the lowest value in a given instance type indicates that the corresponding method performed the most successfully. 
The averages are calculated over all instances of given $V$ and $D$.
Experiments with results in Tab. \ref{tab:small_inst} and Tab. \ref{tab:metamod} are carried out on the same sets of instances.
}

\textcolor{blue}{
The resulting values reveal that the metaheuristic algorithm converges quickly, and is often able to find optimal solutions faster than solving the IP models.
The solutions are, however, not guaranteed to be optimal.
The optimality is proven only on smaller instances, where the B\&B finds the optimum before the time limit is exceeded.
It is likely that the inexact method run on large instances does not find the optimum, but it very clearly outperforms B\&B, 
where the solutions found within the given time limit are far from the optimum.
}

@@

We compare the progress of the lower bounds obtained by the CG method with the one obtained by B\&B applied to $\mathcal{F}_1$ as a function of running time.
Instances with $|V|=2|D|\in\{24, 26,\ldots, 34\}$ are investigated, with 5 instances for each value of $|V|$.
The results are illustrated in Fig. \ref{fig:cggr}.
Each curve represents the progress of the lower bound within each instance set.
The horizontal axes represent running times ranging from 0 to 20 CPU-minutes.
The vertical axes correspond to the lower bounds as a percentage of $z\left(\text{LP}(\mathcal{F}_1)\right)$, averaged over all instances in the set.
\textcolor{blue}{The values can be retrieved from the output of Cplex during the course of simplex and B\&B algorithm.}

@@

    \begin{subfigure}[b]{0.49\textwidth}
        \includegraphics[width=\textwidth]{lower-bound-32-16}
        \caption{$|V|=32, |D|=16$}
        \label{fig:cggr32-16}
    \end{subfigure}
    \hfill %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
      %(or a blank line to force the subfigure onto a new line)
    \begin{subfigure}[b]{0.49\textwidth}
        \includegraphics[width=\textwidth]{lower-bound-34-17}
        \caption{$|V|=34, |D|=17$}
        \label{fig:cggr34-17}
    \end{subfigure}
  \caption{Comparison of lower bounds obtained by CG and B\&B on model $\mathcal{F}_1$.
	\textcolor{blue}{Each figure depicts a progress of lower bounds for instances with varying number of destination and non-destination nodes.
	The lower bounds are obtained by solving CG scheme applied on model $\mathcal{F}_1$ (black curves), and produced during the course of classic B\&B algorithm (grey curves)
	The tightness of the bounds increases with time, and it can be observed from the graphs that the CG bound becomes tighter at some point.
	We express the quality of lower bounds with respect to the value of simple continuous relaxation, as indicated by the vertical axes.
	It can also be noted that the lower bound obtained from B\&B grows sharply during the computation of the simplex method in the root of the B\&B tree, 
	which is happening until the curve reaches 100\%.
	After that point the bound grows rather slowly.
	CG on the other hand produces bound that grows faster even after the value of first relaxation in the root is reached.
%	The values can be retrieved from the output of CPLEX during the course of simplex and B\&B algorithm.
	}}
  \label{fig:cggr}
\end{figure}
